Inverse Modular Reinforcement Learning on Human Motion

### 2015-02-11 paper.tex
added basic content

\title{Inverse Modular Reinforcement Learning on Human Behavior}

\author{
Shun Zhang\\
Department of Computer Science\\
University of Texas at Austin\\
Austin, TX 78712 \\
\texttt{menie482@cs.utexas.edu} \\
\And
Matthew Tong \\
Affiliation \\
Address \\
\texttt{email} \\
\And
Mary Hayhoe \\
Affiliation \\
Address \\
\texttt{email} \\
\And
Dana Ballard \\
Affiliation \\
Address \\
\texttt{email} \\
\\
}

@@

\begin{abstract}
The \emph{title} should be a maximum of 100 characters.

The \emph{abstract} should be a maximum of 2000 characters of text,
including spaces (no figure is allowed). You will be asked to copy
this into a text-only box; and it will appear as such in the
conference booklet. Use 11~point type, with a vertical spacing of
12~points.  The word \textbf{Abstract} must be centered, bold, and in
point size 12. Two line spaces precede the abstract.
\end{abstract}

@@

\acknowledgements{We are deeply indebted to iCORE for
their generous support of RLDM2015.}

@@

@@

A human has not done this kind of task before can achieve a good performance
easily, shown in the experiments described later. If a human knows the policies
of the sub-tasks, or sub-MDPs, he can accomplish a complicated behavior by
combining the sub-MDPs. That is,
$$Q(s, a) = \sum_i w_i Q_i (s, a)$$
where $Q_i$ is the Q value of the i-th sub-MDP, $w_i$ is the weight of the i-th
sub-MDP. $w_i \geq 0, \sum_i w_i = 1$.

% scheduling

\begin{figure}[h!]
\includegraphics[width=0.5\textwidth]{task_1_room_2.png}
\includegraphics[width=0.5\textwidth]{task_4_room_25.png}
\caption{}
\label{fig:exp}
\end{figure}

@@

From a different perspective, can we find a weight vector to best interpret
human's behavior? In Figure~\ref{fig:exp}, there are two scenarios. Same as
Figure~\ref{fig:avatar}, the red circles are obstacles. The blue circles are
targets. The gray line is the path. The black lines are trajectories of human
data. The left figure shows the case where human follows the path and ignores
the targets and obstacles. The right figure shows the case that human does three
sub-tasks simultaneously. The human data are collected by the Center for
Perceptual Systems in University of Texas at Austin.

@@

>>>>

\title{Inverse Modular Reinforcement Learning on Human Motion}

\author{
Shun Zhang\\
Department of Computer Science\\
University of Texas at Austin\\
Austin, TX 78712 \\
\texttt{menie482@cs.utexas.edu} \\
\And
Matthew Tong \\
Center for Perceptual Systems\\
University of Texas at Austin\\
Austin, TX 78712 \\
\texttt{mhtong@gmail.com} \\
\AND
Mary Hayhoe \\
Center for Perceptual Systems\\
University of Texas at Austin\\
Austin, TX 78712 \\
\texttt{hayhoe@utexas.edu} \\
\And
Dana Ballard \\
Department of Computer Science\\
University of Texas at Austin\\
Austin, TX 78712 \\
\texttt{dana@cs.utexas.edu} \\
\\
}

@@

\begin{abstract}
%The \emph{abstract} should be a maximum of 2000 characters of text, including
%spaces (no figure is allowed).
\end{abstract}

@@

\acknowledgements{}

@@

Human is able to learn complicated behaviors much faster than machines can do.

We analyze human's behavior of accomplishing various tasks. With the best of our
knowledge, there is no work using inverse reinforcement learning to understand
human motional behavior in the literature.

\begin{figure}[h!]
\centering
\includegraphics[width=0.6\textwidth]{avatar.png}
\caption{The task of collecting targets, avoiding obstacles and following the
path.}
\label{fig:avatar}
\end{figure}

@@

A human has not done this kind of task before can achieve a good performance
easily, shown in the experiments described later.

\section{Modular Reinforcement Learning}

Value functions is decomposed \cite{koller1999computing}.

 If a human knows the policies
of the sub-tasks, or sub-MDPs, he can accomplish a complicated behavior by
combining the sub-MDPs. That is,
$$Q(s, a) = \sum_i w_i Q_i (s, a)$$
where $Q_i$ is the Q value of the i-th sub-MDP, $w_i$ is the weight of the i-th
sub-MDP. $w_i \geq 0, \sum_i w_i = 1$.

@@

\section{Experiments}

We conducted experiments to ask volunteers to accomplish different tasks and
recorded their trajectories. There are four kind of tasks. Task 1, following the
path only, and ignoring other objects. Task 2, following the path, while avoid
the obstacles. Task 3, following the path, while attain targets when possible.
Task 4, following the path, collecting the targets and avoiding obstacles
simutanously.

\begin{figure}[h!]
\centering
\includegraphics[width=0.24\textwidth]{task_1.png}
\includegraphics[width=0.24\textwidth]{task_2.png}
\includegraphics[width=0.24\textwidth]{task_3.png}
\includegraphics[width=0.24\textwidth]{task_4.png}
\caption{}
\label{fig:exp}
\end{figure}

From a different perspective, can we find a weight vector to best interpret
human's behavior. In Figure~\ref{fig:exp}, same as Figure~\ref{fig:avatar}, the
red circles are obstacles. The blue circles are targets. The gray line is the
path. The black lines are trajectories of human . The left figure shows the case
where human follows the path and ignores the targets and obstacles. The right
figure shows the case that human does three sub-tasks simultaneously. The human
data are collected by the Center for Perceptual Systems in University of Texas
at Austin.

We make some constraints on our learning agent to make it behavior like humans.
It is allowed to go straight ahead, turn left with a small step, or turn right
with a small step.

@@

\begin{figure}[h!]
\centering
\includegraphics[width=0.4\textwidth]{objValuesTask1.png}
\includegraphics[width=0.4\textwidth]{objValuesTask2.png}
\includegraphics[width=0.4\textwidth]{objValuesTask3.png}
\includegraphics[width=0.4\textwidth]{objValuesTask4.png}
\caption{Heatmaps of the objective values of different weights for the four
tasks, respectively. The red zones indicate higher values. The upper two are
Task 1 and 2. The lower two are Task 3 and 4.}
\label{fig:heatmap}
\end{figure}

\begin{table}[h]
\centering
\begin{tabular}{| l| l| l |}
\hline
Average by Task & Num Targs Hit & Num Obst Hit\\
\hline
1 & 2.34 &  2.13\\
\hline
2 & 3.03 &  0.13\\
\hline
3 & 10.19 & 2.28\\
\hline
4 & 9.88 &  0.03\\
\hline
\end{tabular}
\caption{Number of targets hit and number of obstacles hit for human's behavior.}
\end{table}

In Figure~\ref{fig:heatmap}, we show the object values for different weights.

\section{Conclusion and Future Work}

Weighted sum of Q function is one way to combine multiple sub-MDPs. We also
propose other ways, for example, scheduling between different modules, with only
one active at one time. However, when a human tries to collect targets while
avoiding obstacles, these two modules should be both on. A scheduling approach
may yield frequent oscination between target and obstacle modules.
